---
title: "FinalProjectEDA"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)
library(DataExplorer)
```

# Read in Data

```{r}
dat <- read_csv("C:/Users/jacks/Box/BIOS719_G3/Project2/CD4_Botswana.csv")
```

# Explore

```{r}
plot_missing(dat) #cd4 has 22.74% missing data values
plot_histogram(dat)
```

```{r bmi}
summary(dat$bmi)
#The maximum bmi is 130.14 
#obese is 30-39.9
bmi_over_100 <- dat %>% filter(bmi > 100)
# Even though there is a consistent measurement at 130.14, I think it should be removed
# because it is just so large
dat %>% filter(bmi < 100) %>% summary()
# a value of 60 has been deemed reasonable for bmi
table(dat$bmicat) #looks good
```

```{r visit_bid}
table(dat$visit) #looks good
length(unique(dat$bid)) #455 unique bid
```

```{r cd4}
boxplot(dat$cd4) #looks fine to me
boxplot(dat$basecd4) #one definite outlier and a possible other one
dat %>% filter(basecd4 >1500) #one patient at 3663 and one at 1546 (17 rows)
table(dat$cd4cat) #looks good
```

```{r rna}
#this is viral load at baseline
boxplot(dat$baserna)
summary(dat$baserna)
plot(dat$baserna) #seems to be a pretty clear line at about 600,000, and there are
# 41 rows with a value of 750000 which is strange, mostly different bids too
table(dat$rnacat) #looks good
```

```{r anemia}
summary(dat$anemia) #119 NA
table(dat$anemia)
```

```{r sex}
table(dat$sex)
```

```{r}
summary(dat$age)
```

```{r}
table(dat$y2)
```


```{r CleanData}
dat_cleaned <- dat %>% 
  filter(bmi < 100,
         basecd4 < 2000,
         baserna < 600000) %>% 
  mutate(visit2 = visit^2,
         cd4log = log(cd4))

# dat_cleaned$cd4diff <- dat_cleaned$cd4 - dat_cleaned$basecd4

```


```{r}
library(lme4)

fit7 <- lmer(cd4log ~ visit +visit2+ as.factor(cd4cat) + bmi + visit*age + as.factor(cd4cat)*bmi + as.factor(cd4cat)*sex+ as.factor(rnacat)*anemia +visit*as.factor(rnacat) + (1 + visit+ visit*as.factor(cd4cat)|bid), data =dat_cleaned)

summary(fit7)
```

```{r}
dat_to_pred1 <- data.frame(visit = 1, visit2 = 1, cd4cat = 1, bmi = 21.96712, age = 26, sex = 1,rnacat = 1, anemia = 1, bid = "B002486-8")
predicted1 <- predict(fit7, newdata = dat_to_pred1)
cd1t1 <- exp(predicted1)

dat_to_pred12 <- data.frame(visit = 3, visit2 = 9, cd4cat = 1, bmi = 21.96712, age = 26, sex = 1,rnacat = 1, anemia = 1, bid = "B002486-8")
predicted12 <- predict(fit7, newdata = dat_to_pred12)
cd1t2 <- exp(predicted12)

cd1t1 - cd1t2
```

```{r}
dat_to_pred2 <- data.frame(visit = 1, visit2 = 1, cd4cat = 2, bmi = 21.96712, age = 26, sex = 1,rnacat = 1, anemia = 1, bid = "B002486-8")
predicted2 <- predict(fit7, newdata = dat_to_pred2)
cd2t1 <- exp(predicted2)

dat_to_pred22 <- data.frame(visit = 3, visit2 = 9, cd4cat = 2, bmi = 21.96712, age = 26, sex = 1,rnacat = 1, anemia = 1, bid = "B002486-8")
predicted22 <- predict(fit7, newdata = dat_to_pred22)
cd2t2 <- exp(predicted22)

cd2t1 - cd2t2
```

```{r}
unique(as.numeric(row.names(ranef(fit7)$bid)))
```




```{r}
ggplot(dat_cleaned, aes(x = visit, y = cd4))+
  geom_smooth(aes(group = cd4cat, color = as.factor(cd4cat)))+
  geom_point(aes(color = as.factor(cd4cat)))
```


# Aim 1

To study trajectories of CD4 count over time by baseline CD4 category (CD4≤350, 350 –
500, and >500) and identify key baseline characteristics associated with the CD4 trend.

__Analysis Plan__

Consider a model utilizing the longitudinal measurements of CD4 counts over time.
Account for a possible correlation of outcomes across visits within a patient. Estimate temporal trends of CD4 counts by baseline CD4 category and assess associations between other baseline covariates and CD4 counts.

```{r}
library(lme4) #for linear mixed effects model
```

```{r}
#Data has 5445 observations
fit1a <- lmer(cd4 ~ visit + (1|bid) + (0 + visit|bid) + (0 + baserna|bid), data = dat_cleaned, REML = TRUE)
summary(fit1a)
```

```{r}
fit1b <- lmer(cd4 ~ visit + (1 + visit | bid), data = dat_cleaned)
summary(fit1b)
```

```{r}
dat_transformed <- dat_cleaned %>%
  mutate(log_cd4 = log(cd4),
         age_sq = age^2)
```



```{r}
fit1c <- lmer(log_cd4 ~ visit + age_sq + visit*basecd4 + (1 + visit + cd4cat | bid), data = dat_transformed)
summary(fit1c)

library(car)
Anova(fit1c)
```




# Aim 2

To identify risk factors of AIDS progression which is defined as eligibility of ART or death due to HIV related disease. 

__Analysis Plan__

Fit a proper model to identify risk factors of AIDS progression [y2]

```{r}
levels(as.factor(dat_cleaned$y2)) #binary outcome
```

```{r}
fit2a <- lmer(y2 ~ visit + (cd4diff), data = dat_cleaned)
summary(fit2a)
```

```{r}
last_visit_dat <- dat_cleaned %>% 
  group_by(bid) %>% 
  arrange(visit) %>% 
  slice(n())
```


```{r}
fit2b <- glm(y2 ~ basecd4 + baserna, family = binomial(link = "logit"), data = last_visit_dat)
summary(fit2b)
```





```{r}
ggplot(dat_cleaned, aes(x = visit, y = cd4diff))+
  geom_smooth(aes(group = cd4cat, color = as.factor(cd4cat)))+
  labs(x = "Visit",
       y = "Difference of CD4 and Baseline CD4",
       title = "CD4 Difference Over Time By Baseline CD4 Group",
       color = "Baseline CD4 Group")+
  theme(panel.grid.minor = element_blank())

sex_labs <- c("Female", "Male")
names(sex_labs) <- c(1, 2)

ggplot(dat_cleaned, aes(x = visit, y = cd4diff))+
  geom_smooth(aes(group = cd4cat, color = as.factor(cd4cat)))+
  labs(x = "Visit",
       y = "Difference of CD4 and Baseline CD4",
       title = "CD4 Difference Over Time By Baseline CD4 Group",
       color = "Baseline CD4 Group")+
  theme(panel.grid.minor = element_blank())+
  facet_grid(~sex, labeller = labeller(sex = sex_labs))

ggplot(dat_cleaned, aes(x = visit, y = cd4diff))+
  geom_point(aes(color = as.factor(cd4cat)), alpha = .3)+
  theme(panel.grid.minor = element_blank())

dat_cleaned1_100 <- dat_cleaned[1:150,]
ggplot(dat_cleaned1_100, aes(x = visit, y = cd4diff))+
  geom_smooth(aes(group = bid, color = bid), se = FALSE)+
  facet_grid(~cd4cat)

```

