---
title: "719HW5"
format: html
editor: visual
---

```{r}
library(pander)
```

| Deaths        | 0   | 1-9 | 10-49 | 50-99 | 100-199 | 200+ |
|---------------|-----|-----|-------|-------|---------|------|
| Leukemia      | 13  | 5   | 5     | 3     | 4       | 18   |
| Other Cancers | 378 | 200 | 151   | 47    | 31      | 33   |
| Total         | 391 | 205 | 156   | 50    | 35      | 51   |

: Deaths from Leukemia and Other Cancers

## Creating Dataframe

```{r}
dose <- c("0", "1-9", "10-49", "50-99", "100-199", "200+")
leukemia <- c(13, 5,5,3,4,18)
other <- c(378, 200, 151, 47, 31, 33)
total <- c(391, 205, 156, 50, 35, 51)
numeric_dose <- c(0, 5, 59/2, 149/2, 299/2, 250)

df <- data.frame(cbind(dose, leukemia, other, total, numeric_dose))
df$leukemia <- as.integer(df$leukemia)
df$other <- as.integer(df$other)
df$total <- as.integer(df$total)
df$numeric_dose <- as.numeric(df$numeric_dose)
df


```

# A)

## Logit

```{r}
fit.logit <- glm(cbind(leukemia, total-leukemia)~numeric_dose, family=binomial(link="logit"), data=df)

```

## Probit

```{r}
fit.probit <- glm(cbind(leukemia, total-leukemia)~numeric_dose, family=binomial(link="probit"), data=df)
```

## Cloglog

```{r}
fit.cloglog <- glm(cbind(leukemia, total-leukemia)~numeric_dose, family=binomial(link="cloglog"), data=df)

```

## Comparison

```{r}
observed_vals <- df["leukemia"]
logit_vals <- round(df[,"total"]*fit.logit$fitted.values, digits=2)
probit_vals <- round(df[,"total"]*fit.probit$fitted.values, digits=2)
cloglog_vals <- round(df[,"total"]*fit.cloglog$fitted.values, digits=2)
```

| Observed              | Logit             | Probit             | Cloglog             |
|-----------------------|-------------------|--------------------|---------------------|
| `r observed_vals[1,]` | `r logit_vals[1]` | `r probit_vals[1]` | `r cloglog_vals[1]` |
| `r observed_vals[2,]` | `r logit_vals[2]` | `r probit_vals[2]` | `r cloglog_vals[2]` |
| `r observed_vals[3,]` | `r logit_vals[3]` | `r probit_vals[3]` | `r cloglog_vals[3]` |
| `r observed_vals[4,]` | `r logit_vals[4]` | `r probit_vals[4]` | `r cloglog_vals[4]` |
| `r observed_vals[5,]` | `r logit_vals[5]` | `r probit_vals[5]` | `r cloglog_vals[5]` |
| `r observed_vals[6,]` | `r logit_vals[6]` | `r probit_vals[6]` | `r cloglog_vals[6]` |

```{r}
logit_sum <- summary(fit.logit)
logit_coef <- logit_sum$coefficients

probit_sum <- summary(fit.probit)
probit_coef <- probit_sum$coefficients

cloglog_sum <- summary(fit.cloglog)
cloglog_coef <- cloglog_sum$coefficients
```

| Link Function | $\hat{\beta_0}$                | SE($\hat{\beta_0}$)            | $\hat{\beta_1}$                | SE($\hat{\beta_1}$)            | Null Deviance                     | Residual Deviance                 | AIC                          |
|---------------|--------------------------------|--------------------------------|--------------------------------|--------------------------------|-----------------------------------|-----------------------------------|------------------------------|
| Logit         | `r round(logit_coef[1,1],3)`   | `r round(logit_coef[1,2],3)`   | `r round(logit_coef[2,1],3)`   | `r round(logit_coef[2,2],3)`   | `r round(logit_sum$deviance,3)`   | `r round(logit_sum$deviance,3)`   | `r round(logit_sum$aic,3)`   |
| Probit        | `r round(probit_coef[1,1],3)`  | `r round(probit_coef[1,2],3)`  | `r round(probit_coef[2,1],3)`  | `r round(probit_coef[2,2],3)`  | `r round(probit_sum$deviance,3)`  | `r round(probit_sum$deviance,3)`  | `r round(probit_sum$aic,3)`  |
| Cloglog       | `r round(cloglog_coef[1,1],3)` | `r round(cloglog_coef[1,2],3)` | `r round(cloglog_coef[2,1],3)` | `r round(cloglog_coef[2,2],3)` | `r round(cloglog_sum$deviance,3)` | `r round(cloglog_sum$deviance,3)` | `r round(cloglog_sum$aic,3)` |

```{r}
plot(df$numeric_dose, df$leukemia/df$total, pch=19,ylab="Proportion of Leukemia", xlab="Log dosage")
lines(df$numeric_dose, fitted(fit.logit))
lines(df$numeric_dose, fitted(fit.probit), lty=2)
lines(df$numeric_dose, fitted(fit.cloglog), lty=3)
legend("topleft", c("Logit", "Probit", "Cloglog"), lty=c(1,2,3))
```

# B)

The model that fits the data best is the cloglog model, but the logit model is very similar in AIC, standard error of parameter estimates, and deviance. Since the logit model is easier to interpret, I will select that model as my final and "best" model. This can be somewhat seen in the above plot as well, comparing each of the three fitted models with each other. The probit model is what stands out most, as it is the farthest away from the data points at each point that can be deciphered.

# C)

*Interpret results of best model*

$$
logit(p) = \hat{\beta_0} + \hat{\beta_1} X_1
$$

$$
logit(p) = -3.57 + 0.012X_1
$$

$$
logit(\frac{\text{Leukemia Odds}}{\text{All Cancer Odds}}) = -3.57 + 0.012(\text{dose})
$$

Based off of this model, we can say that a 1 unit dose increase results in a 0.012 increase in $logit(p)$ , which based off of the definition of the logit link function, is a 0.012 increase in $log(\frac{p}{1-p})$ , which is the log odds. This can further be interpreted to mean a $exp(0.012) = 1.012$ increase, or a 1.2% increase in the odds of having leukemia compared to having any form of cancer.
