---
title: "719HW6"
format: html
editor: visual
---

## Question1

### A

The saturated model has 6 parameters. 1 from the intercept, 1 from the treatment, 2 from the disease, and 2 from the interaction of disease and treatment.

### B

Model 1: $logit(p) = \beta_0 + \beta_1 * T$
Model 2: $logit(p) = \beta_0 + \beta_1 * D$
Model 3: $logit(p) = \beta_0 + (\beta_1*T) + (\beta_2 *D) + (\beta_3 * T*D)$

### C

Model 1 is suitable to answer question 1.
Model 2 is suitable to answer question 2.
Model 3 is suitable to answer question 3.

### D

#### Build df

```{r}
y <- c(51, 67, 42, 143, 97, 40)
n <- c(200, 158, 89, 315, 263, 75)
severity <- rep(c(0,1,2), 2)
treatment <- c(rep(0, 3), rep(1, 3))

df <- cbind.data.frame(y, n, severity, treatment)
df
```
#### Fit models

```{r}
mod_1 <- glm(cbind(y, n-y) ~ treatment, data = df, family = binomial(link = "logit"))
mod_2 <- glm(cbind(y, n-y) ~ severity, data = df, family = binomial(link = "logit"))
mod_3 <- glm(cbind(y, n-y) ~ treatment*severity, data = df, family = binomial(link = "logit"))
```

### E

DF of the null deviance: `r mod_1$df.null`
DF of the residual deviance for Model 1: `r mod_1$df.residual`
DF of the residual deviance for Model 2: `r mod_2$df.residual`
DF of the residual deviance for Model 3: `r mod_3$df.residual`

### F

```{r}

```


### G

### H

### I

### J
