---
title: "719HW6"
format: html
editor: visual
editor_options: 
  chunk_output_type: inline
---

## Question 1

### A

The saturated model has 6 parameters. 1 from the intercept, 1 from the treatment, 2 from the disease, and 2 from the interaction of disease and treatment.

### B

Model 1: $logit(p) = \beta_0 + \beta_1 * T$ Model 2: $logit(p) = \beta_0 + \beta_1 * D$ Model 3: $logit(p) = \beta_0 + (\beta_1*T) + (\beta_2 *D) + (\beta_3 * T*D)$

### C

Model 1 is suitable to answer question 1. Model 2 is suitable to answer question 2. Model 3 is suitable to answer question 3.

### D

#### Build df

```{r}
y <- c(51, 67, 42, 143, 97, 40)
n <- c(200, 158, 89, 315, 263, 75)
severity <- rep(c(0,1,2), 2)
treatment <- c(rep(0, 3), rep(1, 3))

df <- cbind.data.frame(y, n, severity, treatment)
df
```

#### Fit models

```{r}
mod_1 <- glm(cbind(y, n-y) ~ treatment, data = df, family = binomial(link = "logit"))
mod_2 <- glm(cbind(y, n-y) ~ severity, data = df, family = binomial(link = "logit"))
mod_3 <- glm(cbind(y, n-y) ~ treatment*severity, data = df, family = binomial(link = "logit"))
```

### E

DF of the null deviance: `r mod_1$df.null` DF of the residual deviance for Model 1: `r mod_1$df.residual` DF of the residual deviance for Model 2: `r mod_2$df.residual` DF of the residual deviance for Model 3: `r mod_3$df.residual`

### F

```{r}
mod1_sum <- summary(mod_1)
```

The p-value for the model's treatment coefficient is `r mod1_sum$coefficients[2,4]`, which is statistically significant. This means the coefficient is significantly different from 0 and we can use the estimator to estimate the estimand. 

The odds of hospitalization increases by a factor of `r round(exp(mod1_sum$coefficients[2,1]), 3)`, or 35%, when an individual is administered Drug B compared to Drug A. Drug A is better for preventing hospitalization.

### G

```{r}
mod2_sum <- summary(mod_2)
```

The p-value for the model's severity coefficient is `r mod2_sum$coefficients[2,4]`, which is statistically significant. This means the coefficient is significantly different from 0 and we can use it to estimate our parameter. It also means that severity is statistically significantly associated with hospitalization.

A 1 unit increase in severity results in a `r round(exp(mod2_sum$coefficients[2,1]), 3)` odds increase of hospitalization when comparing Drug A to Drug B, which is 23.1% increase.

### H

```{r}
mod3_sum <- summary(mod_3)
```


$$
= exp(\frac{1}{\hat{\beta_1} + \hat{\beta_2} + \hat{2\beta_3}}) = exp(\frac{1}{0.7023 + .511 - 2*(.5122)}) = exp(\frac{1}{.1889}) = exp(1) - exp(.1889) =  1.51

$$
$$
= \frac{exp(\hat{\beta_0} + \hat{\beta_2})} {exp(\hat{\beta_0} + \hat{\beta_1} + \hat{2\beta_2} + \hat{2\beta_3})} = exp(-\hat{\beta_1} - \hat{\beta_2} -\hat{2\beta_3}) = exp(-0.7032 -.511 - 2*(-.5122) ) = exp(-.1898) = 0.827 
$$


95% CI: 

$$
exp()
$$

### I

I can use model 3 to answer question 3. The interaction term is significant, which means the treatment effect changes for different levels of severity, and vice versa.

To show this, I will compare the effect of a change in severity in two different situations, one where treatment == 0 and one where treatment == 1.

```{r}
mod3_b <- mod3_sum$coefficients
```


$$
\frac{exp(\beta_0 + \beta1*0 + \beta_2*1 + \beta_3*0*1) } {exp(\beta_0 + \beta1*0 + \beta_2*2 + \beta_3*0*2)} = \frac{0.6205}{1.0343} = 0.5999
$$

$$
\frac{exp(\beta_0 + \beta1*1 + \beta_2*1 + \beta_3*1*1) } {exp(\beta_0 + \beta1*1 + \beta_2*2 + \beta_3*1*2)} = \frac{0.7503}{0.7494} = 1.0012
$$
This shows that the odds ratio between levels of severity changes differently for different levels of disease. This shows that disease severity is an effect modifier.

### J

NO
YES
YES

Only model 1 vs Model 3 & Model 2 vs Model 3 can be compared using the LRT.

```{r}
anova(mod_1, mod_3, test = "LRT")
```

```{r}
anova(mod_2, mod_3, test = "LRT")

```

The larger model is preferred in both cases since $H_0 = \text{Smaller model is preferred}$, $H_a: \text{Larger model is preferred}$ and $p < 0.05$, so the interaction term should be included.


## Question 2

### A

```{r}
y <- c(55, 52, 57, 55, 50, 50)
n <- c(102, 99, 108, 76, 81, 90)
force <- rep(c(40, 150, 350), 2)
newstor <- c(rep("Control", 3), rep("Treatment", 3))

df2 <- cbind.data.frame(y, n, force, newstor)
```

```{r}
fit2.1 <- glm(cbind(y, n-y) ~ newstor, family = binomial(link = "logit"), data = df2)
fit2.1
```

```{r}
fit2.2 <- glm(cbind(y, n-y) ~ newstor*force, family = binomial(link = "logit"), data = df2)
fit2.1
```


### B

### C

### D

