---
title: "HW7"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

# Question 1

```{r Libraries}
library(tidyverse)
```


```{r ReadData}
dat <- read_csv("table9_13.csv")
```

## A

```{r Contglm}
fit1_cont <- glm(cbind(y, n-y) ~ ., data = dat, family = binomial(link = "logit"))
```

```{r Catglm}
dat2 <- dat
dat2$CAR <- as.factor(dat2$CAR)
dat2$AGE <- as.factor(dat2$AGE)
dat2$DIST <- as.factor(dat2$DIST)
fit1_cat <- glm(cbind(y, n-y) ~ ., data = dat2, family = binomial(link = "logit"))
```

```{r LRT}
anova(fit1_cont, fit1_cat, test = "LRT")
```

The continuous model fits the data better since p > .05.

```{r CatglmSum}
sum_fit1cont <- summary(fit1_cont)
sum_fit1cont
```

## B

$$
logit(p) = \beta_0 + \beta_1\text{CAR}+\beta_2\text{AGE}+\beta_3\text{DIST} + \epsilon
$$

$$
logit(p)=-1.667+0.232(\text{CAR})-.210(\text{AGE})+.259(\text{DIST})
$$

A 1-unit increase in Age (changing from age group 1 to age group 2, for example) results in a `r round(sum_fit1cont$coefficients[3,1],4)` increase in the log odds of the outcome, or a `r round(exp(sum_fit1cont$coefficients[3,1]),4)` factor increase in the odds of the outcome. This actually results in a decrease of the odds of the outcome by 19%.

## C

Pattern 1: Car group 4 and age group 3 and district 1
Pattern 2: Car group 4 and age group 1 and district 0

$$
\frac{exp(\beta_0+4\beta_1+3\beta_2+\beta_3)}{exp(\beta_0+4\beta_1+\beta_2)} = exp(2\beta_2+\beta_3) = exp(2(-.210)+.259) = 0.8513
$$
```{r CIforOR}
L <- cbind(c(0,0, 2, 1))
Lb <- t(L)%*%fit1_cont$coef
varLb <- t(L)%*%sum_fit1cont$cov.unscaled%*%L
CILb <- Lb + c(-1, 1)*qnorm(1-.05/2)*sqrt(varLb)
exp(CILb)
```

The odds of the outcome are 0.8513 times greater in pattern 1 than in pattern 2. This actually means the odds of the outcome are smaller in pattern 1 than in pattern 2, by approximately 15%. The confidence interval does not contain 1, so it is fair to say that the true odds ratio is within those bounds, with 95% confidence.

## D

```{r PoissonReg}
fit1_pois <- glm(y ~ CAR + AGE + DIST + offset(log(n)), family = poisson(link = "log"), data = dat)
sum_fit1_pois <- summary(fit1_pois)
sum_fit1_pois
```


$$
log(\mu_i)= \beta_0 + \beta_1(CAR) + \beta_2(AGE) + \beta_3(DIST) +\epsilon
$$

$\beta_2$ is associated with a $exp(-.17674) = 0.8379976$ factor increase in the RR of the outcome when AGE increases by 1. This is better interpreted by a ~18% decrease in the RR.

$\beta_3$ is associated with a $exp(.21865) = 1.244396$ factor increase in the RR of the outcome when DIST increases by 1. This is also interpreted as a 24% increase in the RR.

## E

```{r SumCont}
sum_fit1cont
```

```{r SumPois}
sum_fit1_pois
```

The logistic model seems to fit better, and it is not possible to do a LRT test because the models are not the same type of model, i.e. 

```{r}
anova(fit1_cont, fit1_pois, test = "LRT")
```

## F

```{r}
poi.pred <- fitted.values(fit1_pois)
poi.pred.unit <- poi.pred / dat$n

log.pred <- fitted.values(fit1_cont)
# poi.pred.unit <- log.pred

plot(1:32, dat$y / dat$n, pch=19, xlab="Covariate patterns", ylab="Y", bty="L")

points(1:32, poi.pred.unit, pch=19, col=2)
points(1:32, log.pred, pch = 19, col = 3)

legend("topleft", c("Observed", "Poisson", "Log"), col=c(1,2,3), pch=c(19,19, 19))
```

I observe a general okay fit, but I would not call it necessarily good. The Poisson and Log values are pretty similar. The value at x=29 for observed is 0, which does raise some concern but the value is 0 in the data, so it is not an issue with the calculation.

## G

Pattern 1: Car group 4 and age group 3 and district 1
Pattern 2: Car group 4 and age group 1 and district 0


$$
log(\mu_i)= \beta_0 + \beta_1(CAR) + \beta_2(AGE) + \beta_3(DIST) +\epsilon
$$



$$
\frac{exp(\beta_0 + 4\beta_1 + 3\beta_2 + \beta_3)}{exp(\beta_0 + 4\beta_1 + \beta_2)} = exp(2\beta_2+\beta_3) = exp(2(-.17674) + .21865) = 0.8738645
$$



```{r}
L <- matrix(c(0,0,2,1), nrow=4,byrow=T)
lrr <- t(L) %*% as.matrix(fit1_pois$coef)
rr <- exp(lrr)
covmat <- summary(fit1_pois)$cov.unscaled
var.lrr <- t(L)%*%covmat%*%L
ci.lrr <- c(lrr-(1.96*sqrt(var.lrr)), lrr+(1.96*sqrt(var.lrr)))
ci.rr <- exp(ci.lrr)
c(rr, ci.rr)
```

The relative risk is .87, which means that pattern 1 has 0.87 times higher risk of outcome than pattern 2. This actually means that there is greater risk in pattern 2, by approximately 13%. The Ci does not contain 1, though it is VERY close. But, since it does not contain 1, the above mentioned interpretation is appropriate. The values are different from part c, and this is due to the offset __Verify this__

## H

```{r}
fit1_quasi <- glm(y ~ CAR + AGE + DIST + offset(log(n)), data = dat, family = quasipoisson)
summary(fit1_quasi)
```

The dispersion parameter was taken to be 0.8392, which is less than 1, so we have underdispersion.
